Quality Criteria for SKOS Vocabularies

### 2012-03-22 background.tex

Interoperability between applications has been the major motivation for specifying in total six \textbf{SKOS integrity conditions}~\cite{Miles2005}, each of which is a statement that defines under which circumstances data are consistent with the SKOS data model. An example condition is ``a resource has no more than one value of skos:prefLabel per language tag''. Tools that can check whether these integrity conditions are met for given data are already available: two of the six integrity conditions are defined formally in the OWL representation of SKOS and can therefore be validated by any OWL reasoner \todo{CM}{true?}. For validating a SKOS vocabulary against the integrity conditions, one can use tools such as the online \textbf{PoolParty Thesaurus Consistency Checker}, or the \textbf{Skosify} command line utility, which can validate SKOS vocabularies and also correct some detected quality problems.

@@

Typical application areas of controlled vocabularies are as classification, indexing, autocompletion, query reformulation/expansion, or serving as a glossary. As we discussed in detail in our earlier work [\ref{CM}{iSem paper?}], these application areas have specific requirements with respect to vocabulary features, such as structure, availability, and documentation of conceptual resources. A vocabulary that doesn't fulfill these requirements can have a negative effect on an application's output and would therefore be judged by the application designers and potentially also by users as being a low-quality vocabulary. Missing relations between conceptual resources, for instance, can reduce retrieval recall when a vocabulary is used for query expansion. Missing entry points to concepts in a vocabulary can impede orientation for human users. Inconsistent relations can hinder queries across distinct datasets\todo{cm}{this example is unclear}, to name some examples.

% Argue why we need to formalize them -> because otherwise they will be implemented inconsistently across applications
Thus, in order to help vocabulary and application designers in judging the quality of SKOS vocabularies, we need higher-level quality criteria that go beyond low-level integrity conditions defining data model consistency. Such quality criteria are already well-established and have extensively been discussed in standardized guidelines [\todo{CM}{ISO refs}], manuals [\todo{CM}{hedden book, skosify ref 3}], and scholarly articles [\todo{cm}{e.g., skosify refs 16/17}].
% TODO: bitte hier einfach nur die referenzen auf die wir uns beziehen zusammenfassen und kategorisieren. Schau dir dazu auch nochmals die SKOSify referenzen und die aus eurem iSemantics paper an.
If we want check existing SKOS vocabularies against these criteria, we first need to identify them and determine how to evaluate them given the syntax, structure and semantics of SKOS. Since we want to provide tool support for quality checking, we must formalize these criteria into computable quality checking functions, in order to ensure consistent implementation and interpretation across tools.

>>>>

Interoperability between applications has been the major motivation for specifying in total six \textbf{SKOS integrity conditions}~\cite{Miles2005}, each of which is a statement that defines under which circumstances data are consistent with the SKOS data model. An example condition is ``a resource has no more than one value of \texttt{skos:prefLabel} per language tag''. Tools that can check whether these integrity conditions are met for given data are already available: two of the six integrity conditions are defined formally in the OWL representation of SKOS and can therefore be validated by any OWL reasoner. For validating a SKOS vocabulary against the integrity conditions, one can use tools such as the online \textbf{PoolParty Thesaurus Consistency Checker}, or the \textbf{Skosify} command line utility, which can validate SKOS vocabularies and also correct some detected quality problems.

@@

Typical application areas of controlled vocabularies are as classification, indexing, autocompletion, query reformulation/expansion, or serving as a glossary. As we discussed in detail in our earlier work \cite{Nagy2011}, these application areas have specific requirements with respect to vocabulary features, such as structure, availability, and documentation of conceptual resources. A vocabulary that doesn't fulfill these requirements can have a negative effect on an application's output and would therefore be judged by the application designers and potentially also by users as being a low-quality vocabulary. Missing relations between conceptual resources, for instance, can reduce retrieval recall when a vocabulary is used for query expansion. Missing entry points to concepts in a vocabulary can impede orientation for human users. Proprietary term definitions can hinder queries across distinct datasets, to name some examples.

% Argue why we need to formalize them -> because otherwise they will be implemented inconsistently across applications
Thus, in order to help vocabulary and application designers in judging the quality of SKOS vocabularies, we need higher-level quality criteria that go beyond low-level integrity conditions defining data model consistency. Such quality criteria are already well-established and have extensively been discussed in standardized guidelines \cite{ISO25964-1:2011,Z39.19:2005}, manuals \cite{Hedden2010,Aitchison2000,Harpring2010}, and scholarly articles \cite{Coronado2009,Soergel1997,Soergel2002}.

%Whilst aformentioned literature also deals with online vocabulary publication and interoperability \cite{ISO25964-1:2011,Hedden2010}, covering requirements for publication on the Web of Data (as in, e.g., \cite{Heath2011,Hogan2010,Allemang2011}) goes beyond their scope. 
%On the other hand, publications targeting dataset and ontology development for the Web of Data don't elaborate on implications of the design decisions for controlled vocabularies. Our work seeks to narrow this gap by...

% TODO: bitte hier einfach nur die referenzen auf die wir uns beziehen zusammenfassen und kategorisieren. Schau dir dazu auch nochmals die SKOSify referenzen und die aus eurem iSemantics paper an.

If we want check existing SKOS vocabularies against these criteria, we first need to identify them and determine how to evaluate them given the syntax, structure and semantics of SKOS. Since we want to provide tool support for quality checking, we must formalize these criteria into computable quality checking functions, in order to ensure consistent implementation and interpretation across tools.

### 2012-03-22 introduction.tex

	\item We identified \todo{CM}{XX} \textbf{quality criteria} for SKOS vocabularies by examining existing guidelines and formalized them into computable \textbf{filtering functions} that identify possibly affected resources in a given SKOS vocabulary.

>>>>

	\item We identified 15 \textbf{quality criteria} for SKOS vocabularies by examining existing guidelines and formalized them into computable \textbf{quality checking functions} that identify possibly affected resources in a given SKOS vocabulary.
